---
title             : "Battleships, exp. 1: intermediate results"
shorttitle        : "Battleships"

author: 
  - name          : "First Author"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Postal address"
    email         : "my@email.com"
    role:         # Contributorship roles (e.g., CRediT, https://casrai.org/credit/)
      - Conceptualization
      - Writing - Original Draft Preparation
      - Writing - Review & Editing
  - name          : "Ernst-August Doelle"
    affiliation   : "1,2"
    role:
      - Writing - Review & Editing

affiliation:
  - id            : "1"
    institution   : "Wilhelm-Wundt-University"
  - id            : "2"
    institution   : "Konstanz Business School"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.
  
  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.
  
  One sentence clearly stating the **general problem** being addressed by  this particular study.
  
  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).
  
  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.
  
  One or two sentences to put the results into a more **general context**.
  
  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.
  
  <!-- https://tinyurl.com/ybremelq -->
  
keywords          : "keywords"
wordcount         : "X"

bibliography      : ["r-references.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
---

```{r setup, include = FALSE}
library("papaja")
r_refs("r-references.bib")
library('reticulate')
library('tidyverse')
library('broom')
library('cowplot')
library('MESS') # for AUCs
library('lsr') # for effect sizes
library('pwr') # for power calculations
library('brms') # for mixed effects modeling
library('BayesFactor') # for Bayesian t test
library('jsonlite') #parsing data from sort_trial

```

```{python, echo=FALSE, eval=FALSE}
import json
import pandas as pd
import numpy as np
from os import path as path

def to_csv(filename):
   dfs=[]
   with open(filename+'.txt') as json_file:
      for i,line in enumerate(json_file):
         dfs.append(pd.read_json(line))
   group_df = pd.concat(dfs);
   cols = list(group_df.columns.values);
   cols.remove('hover_log');
   group_df = group_df[cols]
   group_df.to_csv(filename+'.csv', index=False)
   return(group_df)

to_csv('..\\experiments\\Battleships\\data\\batch1\\jatos_results_batch1')
```

```{r load-data, warning=FALSE}
E1.df <- read_csv('../experiments/Battleships/data/batch1/jatos_results_batch1.csv') %>%
  rename(subj_id = participant_number) %>%
   # we accidentally have two subjects 251: 5c360ede943bda0001db4f1b and 6092f8b1c80be84bfe9fc57d.
  # to fix this, I change 6092f8b1c80be84bfe9fc57d to be 2511
  mutate(subj_id = ifelse(PROLIFIC_PID=='6092f8b1c80be84bfe9fc57d',2511,subj_id))%>%
  mutate(subj_id = factor(subj_id)) %>%
 
  # the dash is breaking r
  mutate(test_part = ifelse(test_part=='non-pretend','nonpretend',test_part))
```

```{r subject-comments}
comments <- E1.df %>%
  filter(!is.na(responses) & is.na(test_part)) %>%
  select(subj_id,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
  unnest(comments)%>%
  select(subj_id,worker_comments)%>%
  filter(!is.na(worker_comments) & worker_comments!="") %>%
  mutate(worker_comments = gsub("[\r\n]", "",worker_comments))

debrief <- E1.df %>%
  filter(!is.na(responses) & is.na(test_part)) %>%
  select(subj_id,responses) %>%
  mutate(comments = map(responses, ~fromJSON(.) %>%as.data.frame()))%>%
  unnest(comments)%>%
  select(subj_id,debrief)%>%
  filter(!is.na(debrief) & debrief != "")%>%
  mutate(debrief = gsub("[\r\n]", "",debrief))

```

## Strategy use:

After completing the judgment blocks, participants were asked 

> "Did you have a strategy that you used for pretending you had no hints? How about telling between players with and without hints - did you have a strategy for that?"

```{r results='asis'}
string.list <- debrief%>%pull(debrief)
cat(paste('-', string.list), sep = '\n')
```

## General comments
Then, we asked participants for more general feedback about the task:

> "That's it! Before we thank you, we would appreciate if you could share any thoughts you had about the experiment, or anything we should take into account when analyzing your data."
         
```{r results='asis'}
string.list <- comments%>%pull(worker_comments)
cat(paste('-', string.list), sep = '\n')
```



```{r comprehension-checks}
num_instructions <- E1.df %>%
  group_by(subj_id)%>%
  summarise(pretend_instructions = mean(pretend_instructions),
            nonpretend_instructions = mean(nonpretend_instructions))

```




```{r num_clicks}

E1.click_df <- E1.df %>%
  select(subj_id, 
         test_part, 
         grid_number, 
         num_clicks,
         click_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend' | test_part=='nonpretend');

E1.median_clicks <- E1.click_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(num_clicks=median(num_clicks)) %>%
  spread(test_part,num_clicks);

E1.lucky_pretenders <- E1.click_df%>%filter(num_clicks==7)%>%pull(subj_id)%>%unique()

p<- E1.click_df %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=num_clicks,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='number of clicks', y='number of games')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2);

ggsave('figures/num_clicks.pdf',p)


E1.clicks_per_board <- E1.click_df %>%
  filter(!(subj_id %in% E1.lucky_pretenders))%>%
  filter(genuine_first | test_part=='pretend') %>%
  group_by(test_part,grid_number) %>%
  summarise(num_clicks=mean(num_clicks)) %>%
  pivot_wider(id_cols=grid_number,
              names_from=test_part,
              values_from=num_clicks);

p <- E1.clicks_per_board %>%
  ggplot(aes(x=pretend,y=nonpretend)) +
  geom_abline(intercept=0,slope=1) +
  geom_point()+
  coord_fixed(ratio=1)

ggsave('figures/num_clicks_per_board.pdf',p)


E1.clicks_per_common_board <- E1.click_df %>%
  filter(!(subj_id %in% E1.lucky_pretenders))%>%
  select(-click_log)%>%
  pivot_wider(id_cols=c('subj_id','grid_number'),
              names_from=test_part,
              values_from=num_clicks) %>%
  drop_na()

p <- E1.clicks_per_common_board %>%
  ggplot(aes(x=pretend,y=nonpretend)) +
  geom_abline(intercept=0,slope=1) +
  geom_jitter()+
  coord_fixed(ratio=1)

ggsave('figures/num_clicks_per_common_board.pdf',p)

```

```{r, echo=FALSE}
num_clicks = c();
nsim = 10000
boards = rep(NaN,25)

for (i in 1:nsim) {
  board = rep(c(0),25);
  while (sum(board)<7) {
    board[sample(1:25,1)]=1
  }
  nc = which(board==1)%>%tail(n=1);
  num_clicks = c(num_clicks, nc);
  if (nc<25) {
    board[(nc+1):length(board)]=NaN;
  }
  boards = rbind(boards,board)
}

simulated_hits <- data.frame(click_number=1:25,mean_hit=colMeans(boards,na.rm=T))
```

## Number of clicks

To sink all ships, players had to click on at least 7 and at most 25 squares. A simulated player that clicks randomly had a median number of clicks of `r num_clicks%>%median()`. Among our players, the median number of clicks was `r E1.median_clicks%>%pull(nonpretend)%>%median()` in both pretend and non-pretend games. There was no significant difference in the number of clicks between the two conditions (`r wilcox.test(E1.median_clicks$pretend-E1.median_clicks$nonpretend)%>%apa_print()%>%'$'(statistic)`).

On `r E1.click_df%>%filter(num_clicks==7)%>%nrow()` pretend games from `r E1.lucky_pretenders%>%length()` players, games were completed after 7 clicks only, without ever missing a ship. This never happened in non-pretend games. we assumed these participants did not follow task instructions, and excluded them from all following analyses.

```{r num-clicks, fig.pos = 'H', out.width="100%", fig.cap="Number of clicks in pretend (purple) and non-pretend (green) games as a function of block order (pretend or non-pretend first)."}

knitr::include_graphics("figures/num_clicks.pdf")

```

```{r parse-click-log, echo-FALSE}

E1.click_log <- data.frame(matrix(ncol=10,nrow=0, 
                               dimnames=list(NULL, 
                                             c("subj_id",
                                              "test_part", 
                                              "grid_number", 
                                              "num_clicks",
                                              "i",
                                              "j",
                                              "hit",
                                              "t",
                                              "click_number",
                                              "genuine_first"))))


for (row in 1:nrow(E1.click_df)) {
  
    subject_click_log <- data.frame(fromJSON(E1.click_df[row, ]$click_log)) %>%
    mutate(
      click_number = 1:n(),
      subj_id = E1.click_df[row, ]$subj_id,
      test_part = E1.click_df[row, ]$test_part,
      grid_number = E1.click_df[row, ]$grid_number,
      num_clicks = E1.click_df[row, ]$num_clicks,
      genuine_first = E1.click_df[row, ]$genuine_first
    )
    
    E1.click_log <- rbind(E1.click_log, subject_click_log);
}
    

E1.click_log <- E1.click_log %>%
  relocate(subj_id, .before = i) %>%
  relocate(genuine_first, .before = i) %>% 
  relocate(test_part, .before=i) %>% 
  relocate(grid_number, .before=i) %>% 
  relocate(click_number, .before=i) %>%
  group_by(subj_id,test_part,grid_number) %>%
  mutate(RT=t-lag(t,default=0)) %>%
  filter(!(subj_id %in% E1.lucky_pretenders))


```

```{r total_time, echo=FALSE}

E1.total_time_df <- E1.click_log %>%
 filter(click_number==num_clicks)%>%
  mutate(t=t/1000)

E1.median_time <- E1.total_time_df %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(t=median(t)) %>%
  spread(test_part,t);

# anova
E1.time_anva <- aov(t ~ genuine_first * condition, 
                         data = E1.median_time %>%
                           gather('condition','t',nonpretend,pretend) %>%
                           mutate(condition=as.factor(condition)))

p<- E1.total_time_df %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=t,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='duration in seconds', y='number of games')+
    theme(legend.position=c(0.90,0.85)) +
    facet_wrap(~genuine_first, nrow=2);

ggsave('figures/total_time.pdf',p)

E1.time_per_board <- E1.total_time_df %>%
  filter(genuine_first | test_part=='pretend') %>%
  group_by(test_part,grid_number) %>%
  summarise(t=median(t)) %>%
  pivot_wider(id_cols=grid_number,
              names_from=test_part,
              values_from=t);

p <- E1.time_per_board %>%
  ggplot(aes(x=pretend,y=nonpretend)) +
  geom_abline(intercept=0,slope=1) +
  geom_point()+
  coord_fixed(ratio=1)+
  xlim(c(20,40))+
  ylim(c(20,40))

ggsave('figures/time_per_board.pdf',p)

E1.time_per_common_board <- E1.total_time_df %>%
  pivot_wider(id_cols=c('subj_id','grid_number','genuine_first'),
              names_from=test_part,
              values_from=t) %>%
  drop_na()

p <- E1.time_per_common_board %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend')) %>%
  ggplot(aes(x=pretend,y=nonpretend)) +
  geom_abline(intercept=0,slope=1) +
  geom_point()+
  coord_fixed(ratio=1)+
  xlim(c(0,100))+
  ylim(c(0,100)) +
  facet_wrap(~genuine_first,nrow=2)+
    labs(x='duration in seconds, pretend games', 
         y='duration in seconds, non-pretend games',
         title='common board')

ggsave('figures/time_per_common_board.pdf',p)
```

## Game duration

The median game duration was `r E1.total_time_df%>%filter(test_part=='nonpretend')%>%pull(t)%>%median()%>%printnum()` seconds in the non-pretend condition and `r E1.total_time_df%>%filter(test_part=='pretend')%>%pull(t)%>%median()%>%printnum()` seconds in the pretend condition. Pretend games were significantly longer than non-pretend games `r wilcox.test(E1.median_time$pretend-E1.median_time$nonpretend)%>%apa_print()%>%'$'(statistic)`. This difference showed a significant interaction with the order of conditions (`r apa_print(E1.time_anva%>%summary())$full_result$genuine_first_condition`). Note that non-pretend blocks that were played after pretend blocks are not a good representation of players' natural playing behaviour. However, a difference in total game duration was observed also when comparing non-pretend games from non-pretend-first participants with pretend games from pretend-first participants (`r wilcox.test(E1.median_time%>%filter(genuine_first)%>%pull(nonpretend),E1.median_time%>%filter(!genuine_first)%>%pull(pretend))%>%apa_print()%>%'$'(statistic)`).

Overall, participants played 10 games on 9 different boards. 4 boards were used for non-pretend games, 4 for pretend games, and one board was used for both. When zeroing in on game durations for this common board, we observe a positive correlation across participants (and boards), such that participants that took longer in the non-pretend version also took longer in the pretend version (`r cor.test(E1.time_per_common_board$pretend,E1.time_per_common_board$nonpretend)%>%apa_print()%>%'$'(full_result)`). This correlation held also when focusing on the non-pretend-first group (`r cor.test(E1.time_per_common_board%>%filter(genuine_first)%>%'$'(pretend),E1.time_per_common_board%>%filter(genuine_first)%>%'$'(nonpretend))%>%apa_print()%>%'$'(full_result)`).

When taking the median game duration per board after excluding the non-pretend games of pretend-first participants, we see no correlation between pretend and non-pretend game durations across the nine boards (`r cor.test(E1.time_per_board$pretend,E1.time_per_board$nonpretend)%>%apa_print()%>%'$'(full_result)`). 


```{r total-time, fig.pos = 'H', out.width="100%", fig.cap="Total game duration in pretend (purple) and non-pretend (green) games as a function of block order (pretend or non-pretend first)."}

knitr::include_graphics("figures/total_time.pdf")

```


```{r first_RT, eval=FALSE}

E1.first_RT_df <- E1.click_log %>%
 filter(click_number==1)

p<- E1.first_RT_df%>%ggplot(aes(x=t,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity', bins=100) +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(fill="") +
    xlim(0,20000)

ggsave('figures/firstRT.pdf',p)

```
## Reaction times

```{r RT_by_accuracy}

N_perm <- 1000;
bootstrap_error <- function(x, N_perm) {
  N <- length(x)
  medians = c();
  for (i in 1:N_perm) {
    medians = c(medians,sample(x,replace=TRUE,size=N)%>%median())
  };
  return(sd(medians))
}


E1.click_log <- E1.click_log %>%
 mutate(hit_bin = hit !=0,
        lag3 = lag(hit,3)!=0,
        lag2 = lag(hit,2)!=0,
        lag1 = lag(hit,1)!=0,
        lead1 = lead(hit,1)!=0,
        lead2 = lead(hit,2)!=0,
        lead3 = lead(hit,3)!=0)



E1.RT_by_lag3 <- E1.click_log %>%
  filter(click_number>3) %>%
  group_by(subj_id,test_part,lag3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag3,RT,sep='_') %>%
  mutate(lag3_diff = lag3_TRUE-lag3_FALSE);

E1.RT_by_lag2 <- E1.click_log %>%
  filter(click_number>2) %>%
  group_by(subj_id,test_part,lag2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag2,RT,sep='_') %>%
  mutate(lag2_diff = lag2_TRUE-lag2_FALSE);

E1.RT_by_lag1 <- E1.click_log %>%
  filter(click_number>1) %>%
  group_by(subj_id,test_part,lag1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lag1,RT,sep='_') %>%
  mutate(lag1_diff = lag1_TRUE-lag1_FALSE)

E1.RT_by_hit <- E1.click_log %>%
  filter(click_number>0) %>%
  group_by(subj_id,test_part,hit_bin,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(hit_bin,RT,sep='_') %>%
  mutate(hit_diff = hit_bin_TRUE-hit_bin_FALSE)


E1.RT_by_lead1 <- E1.click_log %>%
  filter(click_number<num_clicks) %>%
  group_by(subj_id,test_part,lead1,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead1,RT,sep='_') %>%
  mutate(lead1_diff = lead1_TRUE-lead1_FALSE);

E1.RT_by_lead2 <- E1.click_log %>%
  filter(click_number<num_clicks-1) %>%
  group_by(subj_id,test_part,lead2,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead2,RT,sep='_') %>%
  mutate(lead2_diff = lead2_TRUE-lead2_FALSE)

E1.RT_by_lead3 <- E1.click_log %>%
  filter(click_number<num_clicks-2) %>%
  group_by(subj_id,test_part,lead3,genuine_first)%>%
  summarise(RT=median(RT))  %>%
  spread(lead3,RT,sep='_') %>%
  mutate(lead3_diff = lead3_TRUE-lead3_FALSE)

E1.RT_by_hit_lags <- E1.RT_by_lag3 %>%
  merge(E1.RT_by_lag2) %>%
  merge(E1.RT_by_lag1) %>%
  merge(E1.RT_by_hit) %>%
  merge(E1.RT_by_lead1) %>%
  merge(E1.RT_by_lead2) %>%
  merge(E1.RT_by_lead3);

E1.RT_by_hit_lags_long <- E1.RT_by_hit_lags %>%
  gather("position","difference",c("lag3_diff","lag2_diff", "lag1_diff", "hit_diff", "lead1_diff", "lead2_diff", "lead3_diff"))%>%
  mutate(position = ifelse(position=="lag3_diff", 3,
                           ifelse(position=="lag2_diff",2,
                                  ifelse(position=="lag1_diff",1,
                                         ifelse(position=="hit_diff",0,
                                                ifelse(position=="lead1_diff",-1,
                                                       ifelse(position=="lead2_diff",-2,-3
                                                              ))))))) %>%
  select(subj_id,test_part,genuine_first,position,difference);

p<- E1.RT_by_hit_lags_long %>% 
  group_by(test_part,position,genuine_first) %>%
  summarise(median_difference=median(difference),
            sem_difference=bootstrap_error(difference,N_perm)) %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=position,color=test_part,fill=test_part,y=median_difference)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=-3:3,labels=c('t-3','t-2','t-1','t','t+1','t+2','t+3'))+
  geom_abline(slope=0,intercept=0,linetype=3)+
  geom_vline(xintercept=0,linetype=3)+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_difference-sem_difference,ymax=median_difference+sem_difference),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='Relative serial position', y='RT difference (ms) between hit or miss on timepoint t')+
  theme_classic()

ggsave('figures/RT_by_lag_acc.pdf',p,width=7)


# p<- E1.RT_by_hit %>% 
#   group_by(test_part,hit_bin,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=hit_bin,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit',x='test part', y='median RT (seconds)')
# 
# ggsave('figures/RT_by_acc.pdf',p)
# 
# 
# E1.RT_by_prev_trial <- E1.click_log %>%
#   filter(click_number>1) %>%
#   group_by(subj_id,test_part,prev_hit, genuine_first)%>%
#   summarise(RT=median(RT))
# 
# E1.RT_by_prev_trial_long<- E1.RT_by_prev_trial %>%
#   spread(prev_hit,RT,sep='_') %>%
#   mutate(hit_diff = prev_hit_FALSE-prev_hit_TRUE)
# 
# p<- E1.RT_by_prev_trial %>% 
#   group_by(test_part,prev_hit,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=prev_hit,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit on previous click',x='test part', y='median RT (seconds)')
# 
# ggsave('figures/RT_by_prev_acc.pdf',p)
# 
# E1.RT_by_next_trial <- E1.click_log %>%
#   filter(click_number<num_clicks) %>%
#   group_by(subj_id,test_part,next_hit, genuine_first)%>%
#   summarise(RT=median(RT))
# 
# E1.RT_by_next_trial_long<- E1.RT_by_next_trial %>%
#   spread(next_hit,RT,sep='_') %>%
#   mutate(hit_diff = next_hit_FALSE-next_hit_TRUE)
# 
# p<- E1.RT_by_next_trial %>% 
#   group_by(test_part,next_hit,genuine_first) %>%
#   summarise(median_RT=median(RT)/1000,
#             sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
#   mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
#   ggplot(aes(x=test_part,fill=next_hit,y=median_RT)) +
#   geom_bar(stat="identity",position="dodge", color='black')+
#   geom_errorbar(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),linetype="solid", color="black", position="dodge")+
#   facet_wrap(~genuine_first,nrow=2)+
#   labs(fill='hit on next click',x='test part', y='median RT (seconds)')


```
In non-pretend games, players were faster by `r E1.RT_by_hit_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%median%>%round()` ms when hitting compared to missing a ship (`r E1.RT_by_hit_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`). This effect was exaggerated in pretend games: players were faster by `r E1.RT_by_hit_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%median%>%round()` ms in hits compared to misses (`r E1.RT_by_hit_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`). 

The opposite effect was observed for a hit on the previous click: players were slower by `r E1.RT_by_prev_trial_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms after hitting a target (`r E1.RT_by_prev_trial_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`). Again, this effect was exaggerated in pretend games, where players were slower by `r E1.RT_by_prev_trial_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms compared to misses (`r E1.RT_by_prev_trial_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`).

Finally, when sorting trials according to whether a ship was hit on the next trial, players were slower by `r E1.RT_by_next_trial_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms on the click preceding a hit (`r E1.RT_by_next_trial_long%>%filter(test_part=='nonpretend' & genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`). This effect was not however preserved in pretend games (median difference: `r E1.RT_by_next_trial_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%median%>%round()%>%abs()` ms; `r E1.RT_by_next_trial_long%>%filter(test_part=='pretend' & !genuine_first)%>%pull(hit_diff)%>%wilcox.test()%>%apa_print()%>%'$'(statistic)`).

```{r RT_by_serial_position}

E1.RT_by_position <- E1.click_log %>%
  group_by(subj_id,test_part,genuine_first,click_number)%>%
  summarise(RT=median(RT)) 

E1.RT_by_position_long<- E1.RT_by_position %>%
  spread(click_number,RT,sep='_') 


p<- E1.RT_by_position %>% 
  group_by(test_part,click_number,genuine_first) %>%
  summarise(median_RT=median(RT)/1000,
            sem_RT=bootstrap_error(RT,N_perm)/1000) %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
  ggplot(aes(x=click_number,color=test_part,fill=test_part,y=median_RT)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=seq(1,20,2),limits=c(1,20))+
  geom_path(size=1)+
  geom_ribbon(aes(ymin=median_RT-sem_RT,ymax=median_RT+sem_RT),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='click number', y='median RT (seconds)')+
  ylim(0,3.5)

ggsave('figures/RT_by_serial_position.pdf',p)

```

```{r hit_by_serial_position}

E1.hit_by_position <- E1.click_log %>%
  group_by(subj_id,test_part,genuine_first,click_number)%>%
  summarise(hit=mean(hit_bin)) 

E1.RT_by_position_long<- E1.RT_by_position %>%
  spread(click_number,RT,sep='_') 


p<- E1.hit_by_position %>% 
  group_by(test_part,click_number,genuine_first) %>%
  summarise(mean_hit=mean(hit),
            se_hit=se(hit)) %>%
  mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'),
         simulated_hits = simulated_hits$mean_hit[click_number])%>%
  ggplot(aes(x=click_number,color=test_part,fill=test_part,y=mean_hit)) +
  scale_fill_manual(values=c("#69b3a2", "#404080")) +
  scale_color_manual(values=c("#69b3a2", "#404080")) +
  scale_x_continuous(breaks=seq(1,20,2),limits=c(1,20))+
  geom_path(size=1)+
  geom_path(aes(y=simulated_hits),colour='black',linetype=3,size=1)+
  geom_ribbon(aes(ymin=mean_hit-se_hit,ymax=mean_hit+se_hit),alpha=0.3,show.legend=F)+
  facet_wrap(~genuine_first,nrow=2)+
  labs(fill=' ',x='click number', y='hit probability', colour='')+
  ylim(0,1)

ggsave('figures/hit_probablity_by_serial_position.pdf',p)

```
```{r game_board_analysis, echo=FALSE}

all_possible_boards <- read.csv('../analysis/flat_boards.csv',header=FALSE);

create_board_states <- function(flat_positions,hit) {
  
  board_state = rep(NA,25)
  board_states = c();
  for (p in seq_along(flat_positions)) {
    board_states = c(board_states, paste(board_state,collapse=','));
    board_state[flat_positions[p]]=ifelse(hit[p],1,0);
  }
  
  return(board_states)
};

get_likelihood <- function(board_state) {
  
  board_state = scan(text= board_state, what = numeric(), sep="," , quiet = TRUE);
  
  revealed_positions <- which(!is.na(board_state));
  
  boards_to_remove = c();
  
  for (position in revealed_positions) {
    boards_to_remove = union(boards_to_remove, which(all_possible_boards[,position] != board_state[position]))
  }
  
  survived_boards = setdiff(1:nrow(all_possible_boards), boards_to_remove);
  likelihood = all_possible_boards[survived_boards,]%>%colMeans();
  likelihood[revealed_positions]=NA;
  
  return(paste(likelihood,collapse=','))
  
};

get_posterior <- function(likelihood) {
    likelihood = scan(text= likelihood, what = numeric(), sep="," , quiet = TRUE);
    posterior = likelihood/sum(likelihood,na.rm=TRUE); # assuming a uniform prior
    return(paste(posterior,collapse=','))
}
  
get_p_click <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  return(posterior[flat_position])
}

get_p_click_rank <- function(posterior, flat_position) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  rank_posterior = rank(-posterior)
  return(rank_posterior[flat_position])
}

get_posterior_entropy <- function(posterior) {
  posterior <- scan(text= posterior, what = numeric(), sep="," , quiet = TRUE);
  posterior[which(posterior==0)]=NA; #to avoid infinity*0
  entropy <- -sum(unlist(posterior)*(unlist(log(posterior))),na.rm=TRUE);
  return(entropy);
}


E1.click_log_with_boards <- E1.click_log %>%
  select(subj_id,genuine_first,test_part,grid_number,click_number,i,j,RT,hit_bin) %>% 
  mutate(flat_position = 1+i*5+j) %>%
  group_by(subj_id,genuine_first,test_part,grid_number) %>%
  arrange(click_number) %>%
  summarise(click_number = click_number,
            board_state=create_board_states(flat_position,hit_bin),
            flat_position=flat_position,
            i=i,
            j=j,
            RT=RT,
            hit_bin=hit_bin) %>%
  rowwise()%>%
  mutate(likelihood = get_likelihood(board_state),
         posterior = get_posterior(likelihood),
         p_click = get_p_click(posterior,flat_position),
         entropy = get_posterior_entropy(posterior),
         p_click_rank = get_p_click_rank(posterior,flat_position));



```

```{r p_click}


E1.mean_P_click_first_9_clicks <- E1.click_log_with_boards %>%
  filter(click_number<=9)%>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click=mean(p_click));

E1.mean_P_click_rank <- E1.click_log_with_boards %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(p_click_rank=mean(p_click_rank))

p<- E1.mean_P_click_first_9_clicks %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='p(ship in clicked square)', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2);
ggsave('figures/pclick.pdf',p)

p<- E1.mean_P_click_rank %>%
    mutate(genuine_first = ifelse(genuine_first, 'non-pretend then pretend', 'pretend then non-pretend'))%>%
    ggplot(aes(x=p_click_rank,fill=test_part))+
    geom_histogram( color="#e9ecef", alpha=0.6, position = 'identity') +
    scale_fill_manual(values=c("#69b3a2", "#404080")) +
    labs(x='p(ship in clicked square) - rank', y='number of players')+
    theme(legend.position=c(0.10,0.85)) +
    facet_wrap(~genuine_first, nrow=2);

ggsave('figures/pclick_rank.pdf',p)

E1.num_irrational_clicks <- E1.click_log_with_boards %>%
  group_by(subj_id,test_part, genuine_first) %>%
  summarise(nic=sum(p_click==0)/5) 


 E1.RT_p_click_cor <- E1.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(RT,p_click,method='spearman'));
 
 E1.RT_entropy_cor <- E1.click_log_with_boards %>%
     filter(click_number<9 & click_number>1) %>%
     group_by(subj_id,test_part, genuine_first) %>%
     summarise(r=cor(entropy,p_click,method='spearman'))


```

## Strategy analysis

A near-optimal strategy for Battleships is to set the probability of clicking on a square to be equal to the probability that a ship is hiding in that square, assuming a prior probability over ship locations and given current knowledge about revealed squares. We assumed a uniform probability over all `r all_possible_boards%>%nrow()/2` legal board configurations, and for each click, we calculated the posterior probability for each square on the board. We then gave each square a rank, from 1 being the highest probability to 25 being the lowest one, and compared this mean click score in pretend versus non-pretend games. This rank was higher in non-pretend games, suggesting that participants were more optimal in playing for real compared to when pretending (`r E1.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%mutate(diff=nonpretend-pretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`). This difference was significant in the group that pretended second (`r E1.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%filter(genuine_first)%>%mutate(diff=nonpretend-pretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`), and in the group that pretended first (`r E1.mean_P_click_rank%>%spread(test_part,p_click_rank)%>%filter(!genuine_first)%>%mutate(diff=nonpretend-pretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(statistic)`).

In some cases, participants clicked on squares that could not have a ship in them. For example, clicking right under a revealed horizontal ship is irrational because two ships cannot touch sides. The mean number of irrational clicks for participant was `r E1.num_irrational_clicks%>%spread(test_part,nic)%>%pull(nonpretend)%>%mean()%>%printnum()` in non-pretend games and `r E1.num_irrational_clicks%>%spread(test_part,nic)%>%pull(pretend)%>%mean()%>%printnum()` in pretend games. We found no significant difference between these probabilities `r E1.num_irrational_clicks%>%spread(test_part,nic)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(full_result)`. 

In non-pretend trials, players took longer to click on a square if the posterior probability of this square hiding a ship was low (mean within-subject Spearman correlation: `E1.RT_p_click_cor%>%filter(test_part=='nonpretend')%>%pull(r)%>%mean()%>%printnum()`, `r E1.RT_p_click_cor%>%filter(test_part=='nonpretend')%>%pull(r)%>%t.test()%>%apa_print()%>%'$'(full_result)`). A similar negative correlation between RT and p(ship) was observed in pretend games (`r E1.RT_p_click_cor%>%filter(test_part=='pretend')%>%pull(r)%>%mean()%>%printnum()`, `r E1.RT_p_click_cor%>%filter(test_part=='pretend')%>%pull(r)%>%t.test()%>%apa_print()%>%'$'(full_result)`). A difference between the two correlations was not significant (`r E1.RT_p_click_cor%>%spread(test_part,r)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(full_result)`). However, this correlation was significantly stronger in non-pretend games in the group that played fair first (`r E1.RT_p_click_cor%>%spread(test_part,r)%>%mutate(diff=pretend-nonpretend)%>%filter(genuine_first)%>%pull(diff)%>%t.test()`) and significantly stronger in pretend groups in the group that pretended first (`r E1.RT_p_click_cor%>%spread(test_part,r)%>%mutate(diff=pretend-nonpretend)%>%filter(!genuine_first)%>%pull(diff)%>%t.test()`). In other words, this correlation was stronger in the first condition played, regardless if it was pretend or non-pretend.


In non-pretend trials, players took longer to click on a square when the posterior entorpy was low (mean within-subject Spearman correlation: `E1.RT_entropy_cor%>%filter(test_part=='nonpretend')%>%pull(r)%>%mean()%>%printnum()`, `r E1.RT_entropy_cor%>%filter(test_part=='nonpretend')%>%pull(r)%>%t.test()%>%apa_print()%>%'$'(full_result)`). A similar negative correlation between RT and posterior entropy was observed in pretend games (`r E1.RT_entropy_cor%>%filter(test_part=='pretend')%>%pull(r)%>%mean()%>%printnum()`, `r E1.RT_p_click_cor%>%filter(test_part=='pretend')%>%pull(r)%>%t.test()%>%apa_print()%>%'$'(full_result)`). A difference between the two correlations was not significant (`r E1.RT_entropy_cor%>%spread(test_part,r)%>%mutate(diff=pretend-nonpretend)%>%pull(diff)%>%t.test()%>%apa_print()%>%'$'(full_result)`). 

```{r}

E1.judge_df <- E1.df %>%
  filter(test_part=='judge')%>%
  select(subj_id, 
         grid_number, 
         cheat_player,
         genuine_player,
         cheater,
         noncheater,
         decision,
         replay_log,
         genuine_first) %>%
  mutate(click_log = gsub("\'","\"", click_log)) %>%
  filter(test_part=='pretend' | test_part=='nonpretend');


```
# Discussion


\newpage

# References

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

<div id="refs" custom-style="Bibliography"></div>
\endgroup
